---
title: "Data Analytics CA Pair Project - VR Jungian Sandplay"
author: "Charlie Duff, Adam Heaton, Ademide Adenuga"
date: "December, 2022"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    toc: yes
    toc_float: no
    number_sections: yes
    df_print: paged
  word_document:
    toc: yes
    
---
```{r Load Data, echo=FALSE}
file_name = "data/daie_ca3_data_8.csv"
patient_data <- read.csv(file_name)

patient_data$pre_trial_cpss[is.na(patient_data$pre_trial_cpss)] <- mean(patient_data$pre_trial_cpss,na.rm = TRUE)


```

# Abstract

## Aim and Rationale
The aim and rationale of this project is to formulate a single hypothesis based on a set of data we were given on an experiment conducted to attempt to determine the effectiveness of use of VR technology in Jungian Sandbox therapy, compare the effectiveness of the approaches based on the data we were provided, and cleaning up any data that requires cleanup. 

## Participants and setting
Participants of this experiment were young adults aged between 18 - 25 years old with no exact ages of the participants recorded. 150 participants were recorded in this study with an equal amount of male and female participants recorded. (Male = 75 / Female = 75)
Participants were broken into 3 groups using random sampling.These groups were Control (traditional CBT, w/ No VR), Static (Non-animated model content, w/ VR), Animated (Animated model content, w/ VR)

## Experiment design

## Results gathering
Results of the experiment were tracked in an excel spreadsheet detailing the gender and test group of each participant(Control,Static,Animated) along with their CPSS (Child PTSD Symptom Scale) and OR (Observer Rating) pre and post treatment.

## Major findings


## Findings/Implecations

# Introduction

## Topics And content

## Rationale
The Rationale behind this experiment was to determine if using VR technology in a Jungian Sandbox setting could be beneficial in reducing the PTSD levels in patients. 

# Method
Upon first analysis of the data, a missing datum was found for participant 100 in their Pre-Trial Self Reported CPSS. Between the group, we discussed our options. Between ourselves, two conflicting ideas emerged - one, that we ignore the missing data and calculate results without it (deletion), and the other that we replace the missing data with a suitable value and proceed with an adjusted dataset (imputation). We decided to proceed with imputation using the mean of the column that the missing dataset was relevant to.

First we analysed the missing datum to be Missing Completely at Random (MCAR), as being the sole unrecorded datum in the set, we can conclude that it being missing has nothing to do with any other observed variables (data Missing at Random), or missing due to the values themselves (Missing Not at Random). 

We then discussed how we could impute the missing datum. We came up with a variety of potential options based on research and discussion:
1. Take the mean of all Pre-Trial Self Reported CPSS values. 
2. Take the median of all Pre-Trial Self Reported CPSS values.
3. Take the mode of all Pre-Trial Self Reported CPSS values.
4. Take the mean of all female control Pre-Trial Self Reported CPSS values.
5. Take the median of all female control Pre-Trial Self Reported CPSS values.

The following values were calculated for each option:
1. 6.06
2. 6.09
3. 6.24
4. 5.76
5. 5.82

## Hypothesis
Our hypothesis was that imputing data would provide results with a more varied outcome than data that was ignored.

## Participants
Participants of this experiment were young adults aged between 18 - 25 years old with no exact ages of the participants recorded. 150 participants were recorded in this study with an equal amount of male and female participants recorded. (Male = 75 / Female = 75)
Participants were broken into 3 groups using random sampling. These groups were Control (traditional CBT, w/ No VR), Static (Non-animated model content, w/ VR), Animated (Animated model content, w/ VR)

## Design

## Materials

## Procedure

# Results
```{r table, echo=FALSE}
knitr::kable(patient_data,
  align = c('c'), # a vector with 'c', 'l', or 'r' values
  digits = 2,
  caption = "Patient Data")

pre_trial_cpss_mean <- mean(patient_data$pre_trial_cpss)

```

 
```{r Summary, echo=FALSE}
summary(patient_data)

```
## Graphs
```{r Graphs, echo=FALSE}

plot(density(patient_data$pre_trial_cpss,na.rm = TRUE))
plot(density(patient_data$post_trial_cpss,na.rm = TRUE))
plot(density(patient_data$pre_trial_or,na.rm = TRUE))
plot(density(patient_data$post_trial_or,na.rm = TRUE))


#CPSS Static, Control, Animated
plot(patient_data$pre_trial_cpss[1:50] ~ patient_data$post_trial_cpss[1:50], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Static CPSS", xlab = "Pre-Trial Report", ylab = "Post Trial Report")

plot(patient_data$pre_trial_cpss[51:100] ~ patient_data$post_trial_cpss[51:100], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Control CPSS", xlab = "Pre-Trial Report", ylab = "Post Trial Report")

plot(patient_data$pre_trial_cpss[101:150] ~ patient_data$post_trial_cpss[101:150], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Animated CPSS", xlab = "Pre-Trial Report", ylab = "Post Trial Report")



#OR Static, Control, Animated
plot(patient_data$pre_trial_or[1:50] ~ patient_data$post_trial_or[1:50], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Static OR", xlab = "Pre-Trial Report", ylab = "Post Trial Report")

plot(patient_data$pre_trial_or[51:100] ~ patient_data$post_trial_or[51:100], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Control OR", xlab = "Pre-Trial Report", ylab = "Post Trial Report")

plot(patient_data$pre_trial_or[101:150] ~ patient_data$post_trial_or[101:150], type="p",xlim = c(0,10), ylim = c(0,10), patient_data, main = "Self Reported Data - Animated OR", xlab = "Pre-Trial Report", ylab = "Post Trial Report")

```

## Descriptive statistics
```{r DesStats, include=FALSE}
pre_trial_cpss_meanStatic <- mean(patient_data$pre_trial_cpss[1:50],na.rm = TRUE )
post_trial_cpss_meanStatic <- mean(patient_data$post_trial_cpss[1:50],na.rm = TRUE )

pre_trial_cpss_meanControl <- mean(patient_data$pre_trial_cpss[51:100],na.rm = TRUE )
post_trial_cpss_meanControl <- mean(patient_data$post_trial_cpss[51:100],na.rm = TRUE )

pre_trial_cpss_meanAnimated <- mean(patient_data$pre_trial_cpss[101:150],na.rm = TRUE )
post_trial_cpss_meanAnimated <- mean(patient_data$post_trial_cpss[101:150],na.rm = TRUE )



pre_trial_or_meanStatic <- mean(patient_data$pre_trial_or[1:50],na.rm = TRUE )
post_trial_or_meanStatic <- mean(patient_data$post_trial_or[1:50],na.rm = TRUE )

pre_trial_or_meanControl <- mean(patient_data$pre_trial_or[51:100],na.rm = TRUE )
post_trial_or_meanControl <- mean(patient_data$post_trial_or[51:100],na.rm = TRUE )

pre_trial_or_meanAnimated <- mean(patient_data$pre_trial_or[101:150],na.rm = TRUE )
post_trial_or_meanAnimated <- mean(patient_data$post_trial_or[101:150],na.rm = TRUE )


pre_trial_cpss_std <- sd(patient_data$pre_trial_cpss,na.rm = TRUE )
post_trial_cpss_std <- sd(patient_data$post_trial_cpss)
pre_trial_or_std<- sd(patient_data$pre_trial_or)
post_trial_or_std <- sd(patient_data$post_trial_or)

```
Static Pre trial cpss mean = `r pre_trial_cpss_meanStatic`

Static Post trial cpss mean = `r post_trial_cpss_meanStatic`

Control Pre trial cpss mean = `r pre_trial_cpss_meanControl`

Control Post trial cpss mean = `r post_trial_cpss_meanControl`

Animated Pre trial cpss mean = `r pre_trial_cpss_meanAnimated`

Animated Post trial cpss mean = `r post_trial_cpss_meanAnimated`


Static Pre trial or mean = `r pre_trial_or_meanStatic`

Static Post trial or mean = `r post_trial_or_meanStatic`

Control Pre trial or mean = `r pre_trial_or_meanControl`

Control Post trial or mean = `r post_trial_or_meanControl`

Animated Pre trial or mean = `r pre_trial_or_meanAnimated`

Animated Post trial or mean = `r post_trial_or_meanAnimated`


## Inferential statistics

## Statistical tests
```{r Shapiro-Wilk, include=FALSE}

#perform shapiro-wilk test
result_sw_ptcpss <- shapiro.test(patient_data$pre_trial_cpss)
result_sw_ptcpss
result_sw_postcpss <- shapiro.test(patient_data$post_trial_cpss)
result_sw_postcpss
result_sw_ptor <- shapiro.test(patient_data$pre_trial_or)
result_sw_ptor
result_sw_postor <- shapiro.test(patient_data$post_trial_or)
result_sw_postor


```

A `r result_sw_ptcpss$method` was conducted on each of the data sets

pre trial CPSS data - `r round(result_sw_ptcpss$p.value, 4)`

post trial CPSS data - `r round(result_sw_postcpss$p.value, 4)`

pre trial OR data - `r round(result_sw_ptor$p.value, 4)`

post trial OR data - `r round(result_sw_postor$p.value, 4)`


From the output obtained we can assume normality for the pre and post trial CPSS as the p-value is greater than 0.05 while the OR data fails this normality test.  

```{r Confidence Interval, include=FALSE}

#perform 95% Confidence Interval test
ci_error_precpss <- qt(0.975,df=length(patient_data$pre_trial_cpss)-1)*sd(patient_data$pre_trial_cpss)/sqrt(length(patient_data$pre_trial_cpss))

ci_error_precpss_left <- mean(patient_data$pre_trial_cpss)-ci_error_precpss

ci_error_precpss_right <- mean(patient_data$pre_trial_cpss)+ci_error_precpss


```

95% confidence interval for pre trial CPSS  = (`r ci_error_precpss_left` , `r ci_error_precpss_right` )


## Magnitude and direction of results

# Discussion

# References
Help with getting mean of data while data is missing from column *[Stack Overflow](https://stackoverflow.com/questions/23163863/mean-of-a-column-in-a-data-frame-given-the-columns-name)

Removing Na's by Column *[GeeksforGeeks](https://www.geeksforgeeks.org/remove-rows-with-na-in-one-column-of-r-dataframe/)

How to Handle Missing Data in a Dataset *[freeCodeCamp](https://www.freecodecamp.org/news/how-to-handle-missing-data-in-a-dataset/)

How to Impute Missing Values in R? *[GeeksforGeeks](https://www.geeksforgeeks.org/how-to-impute-missing-values-in-r/)

First 10 entries in a bar plot
*[StackOverflow](https://stackoverflow.com/questions/37524701/how-do-i-barplot-only-the-first-10-entries-in-a-frequency-table-in-r)

Calculate confidence interval *[Cyclismo](https://www.cyclismo.org/tutorial/R/confidence.html)
